---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup-1, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
	warning = FALSE,
  embed_fonts = TRUE,
  fig.fullwidth = TRUE,
  fig.align = "center",
  use_bookdown= TRUE,
  toc= TRUE,
  toc_depth= 4
)
```

```{r libraries-1, include=FALSE}

#### packages needed
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(gridExtra)) install.packages("girdExtra", repos = "http://cran.us.r-project.org")
if(!require(Hmisc)) install.packages("Hmisc", repos = "http://cran.us.r-project.org")
if(!require(sjlabelled)) install.packages("sjlabelled", repos = "http://cran.us.r-project.org")
if(!require(ggthemes)) install.packages("ggthemes", repos = "http://cran.us.r-project.org")
if(!require(ggtext)) install.packages("ggtext", repos = "http://cran.us.r-project.org")
if(!require(ggrepel)) install.packages("ggrepel", repos = "http://cran.us.r-project.org")
if(!require(doBy)) install.packages("doBy", repos = "http://cran.us.r-project.org")
if(!require(sjmisc)) install.packages("sjmisc", repos = "http://cran.us.r-project.org")
if(!require(webshot)) install.packages("webshot", repos = "http://cran.us.r-project.org")
if(!require(scales)) install.packages("scales", repos = "http://cran.us.r-project.org")
if(!require(ggalt)) install.packages("ggalt", repos = "http://cran.us.r-project.org")
if(!require(janitor)) install.packages("janitor", repos = "http://cran.us.r-project.org")
if(!require(bookdown)) install.packages("bookdown", repos = "http://cran.us.r-project.org")
if(!require(htmlwidgets)) install.packages("htmlwidgets", repos = "http://cran.us.r-project.org")
if(!require(widgetframe)) install.packages("widgetframe", repos = "http://cran.us.r-project.org")
if(!require(officer)) install.packages("officer", repos = "http://cran.us.r-project.org")
if(!require(showtext)) install.packages("showtext", repos = "http://cran.us.r-project.org")
if(!require(colorspace)) install.packages("colorspace", repos = "http://cran.us.r-project.org")
if(!require(patchwork)) install.packages("patchwork", repos = "http://cran.us.r-project.org")
if(!require(ggpubr)) install.packages("ggpubr", repos = "http://cran.us.r-project.org")
if(!require(gt)) install.packages("gt", repos = "http://cran.us.r-project.org")
if(!require(htmltools)) install.packages("htmltools", repos = "http://cran.us.r-project.org")
if(!require(yaml)) install.packages("yaml", repos = "http://cran.us.r-project.org")
if(!require(grid)) install.packages("grid", repos = "http://cran.us.r-project.org")
if(!require(see)) install.packages("see", repos = "http://cran.us.r-project.org")

library(see) # colorblind-friendly ‘Okabe-Ito’ palette
library(ggalt) # contingency table
library(tidyverse) # several tools for data manipulation
library(haven) # read SPSS, Stata, and SAS data
library(gridExtra) # arrange multiple grid-based plots on a page
library(Hmisc) # many functions useful for data analysis
library(sjmisc) # useful to set Na's
library(sjlabelled) # get_labels function
library(ggthemes) # themes for ggplot2
library(ggtext) # fancy text
library(ggrepel) # help with the position of the labels on ggplot2 graph
library(doBy) # 'do' something to data stratified 'by' some variables
library(scales) # add scale symbols and colour to axis
library(janitor) # add totals to tables
library(bookdown) # add totals to tables
library(htmlwidgets)
library(widgetframe)
#library(colorspace) # colour blind friendly pallettes 
library(officer) # to add additional formatting to tables 
library(patchwork) # to put graphs side by side
library(ggpubr) # easy-to-use functions for creating and customizing ‘ggplot2’- based publication ready plots.
library(gt) #  creation of interactive data tables
library(htmltools) # tools for creating, manipulating, and writing HTML 
library(grid)
library(webshot)
library(yaml)
library(showtext)
  font_add_google("Lato")
  showtext_auto()

```

```{r load-data-1, include=FALSE}

load("../../data/LFS_clean.rda")

```

```{r copyright-1, include=FALSE}

# top left
copyright = grobTree(textGrob("(c) The Underemployment Project\n      underemployment.info", x=0.008, y= 0.94, hjust=0, gp=gpar(col="gray40", fontsize=8, lineheight = .6, fill="white", alpha = 1))) 

# top right
copyright2 = grobTree(textGrob("(c) The Underemployment Project\n      underemployment.info", x=0.99, y= 0.94, hjust=1, gp=gpar(col="gray40", fontsize=8, lineheight = .6, fill="white", alpha = 1))) 

# bottom left 
copyright3 = grobTree(textGrob("(c) The Underemployment Project\n      underemployment.info", x=0.008, y= 0.05, hjust=0, gp=gpar(col="gray40", fontsize=8, lineheight = .6, fill="white", alpha = 1))) 

```


# Time-related underemployment

<p align="justify">Time-related underemployment refers to individuals who work fewer hours than they desire [@Bell2013; @Wang2018].</p>

<p align="justify">We measure time-related underemployment using three main indicators:</p>
1.	<p align="justify">Part-timers who work part-time because they could not find a full-time job.</p>
2.	<p align="justify">Workers who would like to work longer hours in their current job.</p>
3.	<p align="justify">Workers seeking a replacement job with more hours.</p>

See [Technical Appendix 2: Time-related underemployment] for more details.


## Time-underemployment indicator 1: Involuntary part-time

<p align="justify">People in employment can work either full-time or part-time. A part-time worker is someone who works fewer hours than a full-time worker. In the UK there is no specific number of hours that makes someone full or part-time, and the full-time/part-time threshold is debated, but a full-time worker will usually work 30 or 35 hours or more a week [@GOVND]. In the UK Labour Force Survey, the split between full-time and part-time employment is based on respondents’ self-classification [@ONS2023]. The average hours worked by part-timers, for all the periods included in this analysis, is 20 hours.</p>

<p align="justify">Part-time jobs are of lower quality (including being lower paid) than full-time jobs, on the whole, and they bring a range of career penalties [@ILO2013b; @Nightingale2019; @Warren2018]. Yet part-time employment is extensive in the UK, especially among women, as can be seen in Figure \@ref(fig:plot-involuntary-part-time-by-sex). The Labour Force Survey disaggregates part-time employment by reason as follows:</p>

- **those who could not find full-time jobs**
- those who did not want full-time jobs
- those who were ill or were disabled
- those who were students or were at school

<p align="justify">**We focus on those workers who could not find a full-time job** as our first indicator of underemployment.</p>  


### Time-underemployed part-time workers by sex

<p align="justify"> More female than male workers in the UK are employed part-time. Figure \@ref(fig:plot-involuntary-part-time-by-sex) shows which of those part-timers work part-time because they could not find a full-time job. In absolute numbers, there are more female part-timers in this situation (see bottom figure in Figure \@ref(fig:plot-involuntary-part-time-by-sex)). One contributing factor is that, even after taking into account the gender of workers, underemployment is more common in female- dominated occupations, and less prevalent in male-dominated occupations and industries. The percentage of part-timers unable to find full-time work is significantly lower for women than it is for men: it is around 10% in most periods for women compared with a range from 20% to 40% for men (as shown in the top figure of Figure \@ref(fig:plot-involuntary-part-time-by-sex)). Most female part-timers say they do not want full-time jobs, whereas the majority of male part-timers desire full-time employment but cannot find it. A key factor contributing to this disparity is that women often work part-time to fit around other roles in their lives, including caring for children and others [@Warren2018].</p>

```{r involuntary-part-time-by-sex-table, include=FALSE}

options(digits = 3)

t_ptime <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(FTPTWK == 2) %>% # only part-timers
  group_by(year, sequence, quarter, SEX, YPTJOB) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, SEX) %>% # calculate % per group
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(YPTJOB == 3) %>% # could not find full time only
  as_label(.)


```

```{r plot-involuntary-part-time-by-sex, fig.width = 8, fig.asp = 0.7, out.width = '100%', fig.cap= "Part-time workers who could not find a full-time job by sex"}

p_ptime <- t_ptime %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(SEX))) + 
  geom_line(size = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "gray80") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "gray80") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "gray80") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "gray80") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "gray80") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "gray80") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "gray80") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "gray80") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "gray80") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "gray80") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "gray80") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "gray80") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "gray80") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "gray80") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "gray80") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "gray80") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dark 3") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,1, 0.1), limits = c(0, 1), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "top",
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
    # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = c("Note: Could not find a full-time job as a percentage of all part-time workers in the group")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=1, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.75, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=1, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.75, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .6) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.75, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .6) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.75, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .6) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = ggplot2::margin(0, 0, 0, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(4, 0, 5, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "none", 
        legend.title = element_blank(),
        legend.text = element_text(size = 14),
        panel.spacing = unit(.2, "cm"),
        #legend.key.size = unit(0.5, 'cm'),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 5)))


p_ptime_total <- t_ptime %>% 
  ggplot() + 
  geom_bar(aes(x = sequence, y = count, fill = factor(SEX)), stat="identity", position = "dodge", width=0.8) +  
  geom_vline(xintercept = 4.5, color = "gray80") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "gray80") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "gray80") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "gray80") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "gray80") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "gray80") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "gray80") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "gray80") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "gray80") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "gray80") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "gray80") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "gray80") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "gray80") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "gray80") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "gray80") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "gray80") + # gridline 2021

  # scale and colour
  #scale_fill_discrete_qualitative(palette = "Dark 3") +
  scale_fill_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,1200000, 100000), limits = c(0, 1200000), labels = comma, expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68),
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "Year", 
       y = "Total", 
       #title ="",
       #subtitle  = "",
       #caption = c("Source: The Underemployment Project. Read more at underemployment.info")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=1200000, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 1000000, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=1200000, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 1000000, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=1200000, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 1000000, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=1200000, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 1000000, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = ggplot2::margin(0, 0, 0, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 0, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 14),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 5)))

p_ptime / p_ptime_total

```


### Time-underemployed part-time workers by age group

<p align="justify">Figure \@ref(fig:plot-involuntary-part-time-age) shows the part-time workers who could not find full-time work divided by age groups. Part-timers aged between 18 to 24 years old were most likely to find themselves involuntarily working part-time. Time-underemployment among the young increased following economic recessions (in 2008-9 and 2020) but the ongoing cost of living crisis from 2021 has not yet significantly worsened involuntary part-time work for this age group. Trends over time show similar peaks and troughs for other age groups, albeit with notably lower levels of involuntary part-time employment.</p>

```{r involuntary-part-time-by-age-table, include=FALSE}

options(digits = 3)

t_ptime_age <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  #creating age categories
  mutate(AGE_cat = case_when(AGE %in% 18:24 ~ "Aged 18 to 24",
                             AGE %in% 25:34 ~ "Aged 25 to 34",
                             AGE %in% 35:44 ~ "Aged 35 to 44",
                             AGE %in% 45:54 ~ "Aged 45 to 54",
                             AGE %in% 55:64 ~ "Aged 55 to 64",
                             is.na(AGE) ~ NA)) %>%
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  group_by(year, sequence, quarter, AGE_cat, YPTJOB) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, AGE_cat) %>% # calculate % per group
  # percentage per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(YPTJOB == 3) %>%  
  as_label(.)

```


```{r plot-involuntary-part-time-age, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, out.width = '100%', fig.cap= "Young part-time workers struggle the most to find full-time jobs"}

p_ptime_age  <- t_ptime_age  %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(AGE_cat))) + 
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dark 3") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,0.25, 0.025), limits = c(0, 0.25), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "bottom",
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = "Note: Could not find a full-time job as a percentage of all part-time workers in the age group"
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=0.25, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.20, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=0.25, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.20, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=0.25, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.20, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=0.25, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.20, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_ptime_age

```


### Time-underemployed part-time workers by ethnic group

<p align="justify">Work-time in the UK is shaped by ethnic background. Figure \@ref(fig:plot-involuntary-part-time-ethnicity) shows that part-time workers from minority-ethnic groups, grouped together, are more likely to work part-time involuntarily than are white workers. During the period under study both groups experienced similar trends with increases in underemployment following recessions and decreases during economic recoveries. However, minority-ethnic groups were disproportionately affected during recessions, experiencing a larger increase of involuntary part-time work than white groups after each economic crisis.</p>

```{r involuntary-part-time-by-ethnicity-table, include=FALSE}

LFS_clean <- LFS_clean %>% mutate(ETH2 = case_match(ETH, 
                                        1 ~ "White",
                                        2 ~ "Minority-ethnic", 
                                        3 ~ "Minority-ethnic", 
                                        4 ~ "Minority-ethnic", 
                                        5 ~ "Minority-ethnic",
                                        6 ~ "Minority-ethnic")) %>% 
                                    as_numeric(., keep.labels = TRUE)

t_ptime_eth <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  group_by(year, sequence, quarter, ETH2, YPTJOB) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, ETH2) %>% 
  # percentage per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(YPTJOB == 3) %>%  
  filter(!is.na(ETH2)) %>%  # remove NAs from the ethnicity variable
  as_label(.)


```

```{r plot-involuntary-part-time-ethnicity, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, out.width = '100%', fig.cap= "Ethnic-minority groups more likely to find themselves in involuntary part-time jobs"}

p_ptime_eth  <- t_ptime_eth  %>% 
  ggplot(aes(x = sequence, y = percentage, colour = ETH2)) + 
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021
  #facet_wrap(vars(SEX), nrow = 2) +

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dark 3") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,0.30, 0.025), limits = c(0, 0.30), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "bottom",
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="Ethnic-minority groups more likely to find themselves in involuntary part-time",
       #subtitle  = "",
       caption = "Note: Could not find a full-time job as a percentage of all part-time workers in the ethnic group"
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=0.30, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.25, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=0.30, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.25, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=0.30, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.25, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=.30, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.25, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06),
        legend.spacing = unit(0.2, 'cm')) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_ptime_eth

```


### Time-underemployed part-time workers by qualification level

<p align="justify">Qualification levels are a key factor shaping who is time-related underemployed. Figure \@ref(fig:plot-involuntary-part-time-education) shows higher rates of involuntary part-time working among lower/no qualified part-timers. This suggests that  across all years of our analysis, workers with no qualifications or lower-level qualifications find it more difficult to find full-time employment than do those with higher qualifications.</p>

```{r involuntary-part-time-table-education, include=FALSE}

options(digits = 3)

t_ptime_quali <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(FTPTWK == 2) %>% # only part-timers
  group_by(year, sequence, quarter, quali, YPTJOB) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, quali) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(YPTJOB %in% c("3") & !is.na(quali) & !quali %in% c("7")) %>% # could not find
  filter(YPTJOB == 3) %>% # could not find  
  as_label(.)


```

```{r plot-involuntary-part-time-education, fig.width = 7.5, fig.asp = 0.6, out.width = '100%', fig.fullwidth = TRUE, fig.cap= ": Workers with no formal qualification find it difficult to find full-time work"}

p_ptime_quali <- t_ptime_quali %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(quali))) + 
  geom_line(size = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dark 3") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,1, 0.1), limits = c(0, 1), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), 
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title =" ",
       #subtitle  = "",
       caption = c("Note: Could not find a full-time job as a percentage of part-time workers in the category")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=1, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.80, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=1, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.80, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.80, label = "Covid-19", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.80, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
 guides(col = guide_legend(nrow = 2, byrow = TRUE, override.aes = list(size = 6)))

p_ptime_quali

```


### Time- underemployed part-time workers by occupational group 

<p align="justify">Part-time employment is concentrated in lower waged occupations, and with fewer opportunities to work part-time in more senior positions. Figure \@ref(fig:plot-involuntary-part-time-nsec) shows which occupations feature the most involuntary part-time working among part-timers. Routine occupations (e.g., HGV driver, van driver, cleaner, porter, packer, sewing machinist, messenger, waiter or waitress, or bar staff), semi-routine (e.g., postal worker, machine operative, security guard, caretaker, farm worker, catering assistant, receptionist or sales assistant) and lower supervisory/technical (e.g., motor mechanic, fitter, inspector, plumber, printer, tool maker, electrician, gardener or train driver) stand out with higher rates of time-underemployment than managers/professionals and those in intermediate roles. Economic downturns affected these groups the most too, intensifying their underemployment.</p>

```{r involuntary-part-time-table-NSEC, include=FALSE}

options(digits = 3)

t_ptime_nsec <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(FTPTWK == 2) %>% # only part-timers
  group_by(year, sequence, quarter, NSEC2, YPTJOB) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, NSEC2) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(NSEC2) & !NSEC2 %in% c(3,6)) %>%  # omit no answers and no apply
  filter(YPTJOB == 3) %>% # could not find  
  as_label(.)

```


```{r plot-involuntary-part-time-nsec, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Routine, semi-routine and technical part-timers more likely to be in involuntary part-time work"}

p_ptime_nsec <- t_ptime_nsec %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(NSEC2))) + 
  geom_line(size = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dynamic") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,1, 0.1), limits = c(0, 1), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), 
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = c("Note: Could not find a full-time job as a percentage of part-time workers in the category")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=1, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.80, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=1, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.80, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.80, label = "Covid-19", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.80, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_ptime_nsec


```


###  Time-underemployed part-time workers by contract type

<p align="justify">In recent years, the UK has seen a decline in the number of employees with a so-called ‘standard employment relationships’, characterized by a secure and full-time job contract with their employers. The proliferation of alternative contract types, such as flexi-work, term-time working, raises many questions about the quality of that work and workers’ experiences. Perhaps the most contentious form of work, especially relevant for our focus on temporal underemployment, is the zero-hours contract, in which workers are not guaranteed any hours of work at all from their employer. Zero hours contracts give employers a great deal of flexibility, but there is debate over whether workers opt positively into zero hours jobs for flexibility themselves or must agree to such precarious work in a limited labour market. We show here that part-timers with more precarious contracts are more likely to be in part-time work because they could not find a full-time job than are other workers (Figure \@ref(fig:plot-involuntary-part-time-fled): workers with zero-hours contracts have had the highest levels of involuntary part-time work since 2008.</p>  

```{r involuntary-part-time-table-fled, include=FALSE}

# 7 "On call working" is not presente before 2010

# "Flexi-time" = 1,
# "Annualised hours contract" = 2,
# "Term-time working" = 3,
# "Job sharing" = 4,
# "Condensed or compressed hours" = 5,
# "Zero hours contract" = 6, 
# "On call working" = 7,
# "None of these" = 8

options(digits = 3)

t_ptime_fled <- LFS_clean %>% # summary table
  filter(!sequence %in% c(1, 3, 5, 7, 9, 13, 17, 19, 21, 23, 
                        25, 27, 31, 33, 35, 37, 39, 41, 43,
                        45, 47, 49, 51, 53, 55)) %>%
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(FTPTWK == 2) %>% # only part-timers
  group_by(year, sequence, quarter, FLED, YPTJOB) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, FLED) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(FLED) & !FLED %in% c(5, 7, 8)) %>%  # omit compressed, none, and on call
  filter(YPTJOB == 3) %>% # could not find  
  as_label(.)

```

```{r plot-involuntary-part-time-fled, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Workers in zero hours contracts struggle more to find full-time work"}

p_ptime_fled <- t_ptime_fled %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(FLED))) + 
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dynamic") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,1, 0.1), limits = c(0, 1), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), 
                     labels=c("","","2006","",
                              "","2007","","",
                              "","2008","","",
                              "","2009","","",
                              "","2010","","",
                              "","2011","","",
                              "","2012","","",
                              "","2013","","",
                              "","2014","","",
                              "","2015","","",
                              "","2016","","",
                              "","2017","","",
                              "","2018","","",
                              "","2019","","",
                              "","2020","","",
                              "","2021","","",
                              "","2022","",""),
                     expand = c(0,0)) +

   # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = c("Note: Could not find a full-time job as a percentage of part-time workers in the category")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=1, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.80, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=1, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.80, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.80, label = "Covid-19", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.80, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_ptime_fled

```


### Regional trends in time-underemployment among part-time workers

<p align="justify">Our project provides analysis of underemployment across the UK, but it focuses too on the four cities of Bristol, Glasgow, Nottingham and Salford. Figure \@ref(fig:plot-involuntary-part-time-region) shows the distribution of workers in involuntary part-time employment in regions of England and Scotland that are in the scope of this project plus the rest of England. The North West of England (including Salford) and Scotland (including Glasgow) have the highest proportion of workers working part-time because they could not find a full-time work, while the South West (including Bristol) consistently has the lowest proportions, and this pattern has remained consistent over time . Meanwhile, the East Midlands Region (including Nottingham) saw an increase in involuntary part-time working during 2022.</p>
 
```{r involuntary-parttime-region, include=FALSE}

LFS_clean <- LFS_clean %>% mutate(GOVTOF3 = case_match(GOVTOF2, 
                                            1 ~ "Rest of England",
                                            2 ~ "North West", 
                                            4 ~ "Rest of England",
                                            5 ~ "East Midlands", 
                                            6 ~ "Rest of England", 
                                            7 ~ "Rest of England",
                                            8 ~ "Rest of England", 
                                            9 ~ "Rest of England",
                                            10 ~ "South West",
                                            11 ~ "Wales",
                                            12 ~ "Scotland",
                                            13 ~ "Northern Ireland")) %>% 
                                                as_numeric(., keep.labels = TRUE)

LFS_clean$GOVTOF3 <- set_labels(LFS_clean$GOVTOF3, labels = c(
                                                      "East Midlands (incl. Nottingham)" = 1,
                                                      "North West (incl. Salford)" = 2, 
                                                      "Northern Ireland" = 3,
                                                      "Rest of England" = 4, 
                                                      "Scotland (incl. Glasgow)" = 5, 
                                                      "South West (incl. Bristol)" = 6,
                                                      "Wales" = 7))

options(digits = 2)

t_ptime_region <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(FTPTWK == 2) %>% # only part-timer
  group_by(year, sequence, quarter, GOVTOF3, YPTJOB) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, GOVTOF3) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(GOVTOF3)) %>%  # omit no answers and no apply
  filter(YPTJOB == 3) %>% # could not find  
  select(!YPTJOB) %>% 
  as_label(.)

```


```{r plot-involuntary-part-time-region, fig.width = 7.5, fig.asp = 0.6, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Workers in the North West of England and Scotland struggle more to find full-time work"}

p_ptime_region <- t_ptime_region %>% 
  filter(!GOVTOF3 %in% c("Northern Ireland","Wales")) %>% 
  ggplot(aes(x = sequence, y = percentage, col = factor(GOVTOF3))) + 
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dynamic") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,0.5, 0.05), limits = c(0, 0.5), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), 
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = c("Note: Could not find a full-time job as a percentage of part-time workers in the region")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=0.5, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.42, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=0.5, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.42, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=0.5, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.42, label = "Covid-19", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=0.5, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.42, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 0, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 2, byrow = TRUE, override.aes = list(size = 6)))

p_ptime_region

```


### Industry trends in time-underemployment among part-time workers

<p align="justify">In addition to its concern with underemployment across the whole labour force in the UK, the project focuses upon two industry sectors: Wholesale and retail trade, and Human health and social work. We are examining the extent of diverse and intersecting forms of underemployment there and the lived experiences of workers. Figure \@ref(fig:plot-involuntary-part-time-ind1) shows the distribution of workers in involuntary part-time in six service-related sectors. Involuntary part-time work is particularly prevalent among part-timers employed in the Retail as well as Hospitality (accommodation and food service) sectors, justifying their inclusion in our research going forward. Human health and social work activities show lower levels compared to the two sectors mentioned above, and similar rates with sectors such as Education. Very few part-timers working in finance and insurance are doing so involuntarily.</p>

```{r involuntary-part-time-table-ind1, include=FALSE}

options(digits = 3)

t_ptime_ind1 <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(FTPTWK == 2) %>% # only part-timers
  group_by(year, sequence, quarter, ind1, YPTJOB) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, ind1) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(ind1) & ind1 %in% c(7, 17, 11, 16, 10, 9)) %>%  # omit no answers and no apply
  filter(YPTJOB == 3) %>% # could not find  
  as_label(.)

# Of interest:
  # "Wholesale and retail trade" = 7,
  # "Human health and social work" = 17

```


```{r plot-involuntary-part-time-ind1, fig.width = 7.5, fig.asp = 0.6, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Involuntary part-time is high in the Retail and Hospitality sectors"}

p_ptime_ind1 <- t_ptime_ind1 %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(ind1))) + 
  geom_line(size = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dark 3") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,1, 0.1), limits = c(0, 1), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), 
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = c("Note: Could not find a full-time job as a percentage of part-time workers in the category")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=1, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.80, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=1, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.80, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.80, label = "Covid-19", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.80, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 3, byrow = TRUE, override.aes = list(size = 5)))

p_ptime_ind1

```


## Time-underemployment indicator 2: Wanting longer hours

<p align="justify">Our second indicator of time-related underemployment moves beyond the focus only on part-time workers. We explore next all workers who say that they would like to work longer hours in their jobs (at their current basic rate of pay). The following sections look at this group of workers by gender, age, ethnicity and so on.</p>


### Wanting more hours by sex

<p align="justify">The second indicator of time-related underemployment produces similar findings to the first one (involuntary-part-time). Figure \@ref(fig:plot-longer-hours-sex) shows women and men who wish to work longer hours in their current jobs. Female workers are slightly more likely than are men to say they want more hours. The trends over time follow a similar pattern for male and female workers characterised by periods of increased underemployment during and after economic recessions.</p>


```{r longer-hours-gender-table, include=FALSE}

options(digits = 3)

t_undemp_sex <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 2) %>% # Applies to all respondents who are not looking for a different or additional job
  group_by(year, sequence, quarter, SEX, UNDEMP) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, SEX) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(UNDEMP)) %>% # remove NAs. Only responses to the question
  filter(UNDEMP == 1) %>% # only those wating to work additional hours
  as_label(.)


```


```{r plot-longer-hours-sex, ffig.width = 8, fig.asp = 0.8, out.width = '100%', fig.cap= "Women slightly more represented among workers wanting more hours"}

p_undemp_sex_rate <- t_undemp_sex %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(SEX))) + 
  geom_line(size = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "gray80") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "gray80") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "gray80") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "gray80") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "gray80") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "gray80") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "gray80") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "gray80") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "gray80") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "gray80") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "gray80") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "gray80") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "gray80") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "gray80") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "gray80") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "gray80") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dark 3") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,0.3, 0.025), limits = c(0, 0.3), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "top",
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
    # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       #caption = c("")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=0.3, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.27, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=0.3, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.27, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .6) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=0.3, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.27, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .6) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=0.3, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.27, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .6) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0.4, 0), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = ggplot2::margin(0, 0, 0, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(4, 0, 5, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "none", 
        legend.title = element_blank(),
        legend.text = element_text(size = 14),
        panel.spacing = unit(.2, "cm"),
        #legend.key.size = unit(0.5, 'cm'),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 5)))


p_undemp_sex_total <- t_undemp_sex %>% 
  ggplot() + 
  geom_bar(aes(x = sequence, y = count, fill = factor(SEX)), stat="identity", position = "dodge", width=0.8) +  
  geom_vline(xintercept = 4.5, color = "gray80") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "gray80") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "gray80") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "gray80") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "gray80") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "gray80") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "gray80") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "gray80") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "gray80") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "gray80") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "gray80") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "gray80") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "gray80") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "gray80") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "gray80") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "gray80") + # gridline 2021

  # scale and colour
  #scale_fill_discrete_qualitative(palette = "Dark 3") +
  scale_fill_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,3000000, 300000), limits = c(0, 3000000), labels = comma, expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68),
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "Year", 
       y = "Total", 
       #title ="",
       #subtitle  = "",
       caption = c("Note: Applies to all respondents who are not looking for a different or additional job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=3000000, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 2550000, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=3000000, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 2550000, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=3000000, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 2550000, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=3000000, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 2550000, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0.4, 0), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 0, margin = ggplot2::margin(0, 0, 0, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 0, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 14),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_undemp_sex_rate / p_undemp_sex_total

```


### Wanting more hours by age group

<p align="justify">The second indicator of underemployment again reiterates our earlier finding of higher levels of underemployment among younger workers. Figure \@ref(fig:plot-longer-hours-age) shows that the younger the worker is, the more likely they are to want longer hours in their current jobs. As with involuntary part-time, this form of underemployment peaked after economic recessions, particularly among younger workers.</p> 

```{r longer-hours-age-table, include=FALSE}

options(digits = 3)

t_undemp_age <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  #creating age categories
  mutate(AGE_cat = case_when(AGE %in% 18:24 ~ "Aged 18 to 24",
                             AGE %in% 25:34 ~ "Aged 25 to 34",
                             AGE %in% 35:44 ~ "Aged 35 to 44",
                             AGE %in% 45:54 ~ "Aged 45 to 54",
                             AGE %in% 55:64 ~ "Aged 55 to 64",
                             is.na(AGE) ~ NA)) %>%
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 2) %>% # Applies to all respondents who are not looking for a different or additional job
  group_by(year, sequence, quarter, AGE_cat, UNDEMP) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, AGE_cat) %>% # calculate % per group
  # percentage per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(UNDEMP)) %>% # remove NAs. Only responses to the question
  filter(UNDEMP == 1) %>% # only those wating to work additional hours
  as_label(.)

```

```{r plot-longer-hours-age, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.cap= "Young workers wish to work more hours"}

p_undemp_age  <- t_undemp_age  %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(AGE_cat))) + 
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dark 3") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,0.50, 0.05), limits = c(0, 0.50), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "bottom",
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = "Note: Applies to all respondents who are not looking for a different or additional job"
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=0.5, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.40, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=0.5, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.40, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=0.5, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.40, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=0.5, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.40, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_undemp_age

```


### Wanting more hours by ethnic group

<p align="justify">Figure \@ref(fig:plot-longer-hours-eth) shows preferences for more hours of work by ethnic group. Minority-ethnic workers are again more likely to want more hours than are the white majority each year, and with slightly fluctuating levels over time.</p>

```{r longer-hours-eth-table, include=FALSE}

LFS_clean <- LFS_clean %>% mutate(ETH2 = case_match(ETH, 
                                        1 ~ "White",
                                        2 ~ "Minority-ethnic", 
                                        3 ~ "Minority-ethnic", 
                                        4 ~ "Minority-ethnic", 
                                        5 ~ "Minority-ethnic",
                                        6 ~ "Minority-ethnic")) %>% 
                                    as_numeric(., keep.labels = TRUE)

t_undemp_eth <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 2) %>% # Applies to all respondents who are not looking for a different or additional job
  group_by(year, sequence, quarter, ETH2, UNDEMP) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, ETH2) %>% 
  # percentage per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(UNDEMP)) %>% # remove NAs. Only responses to the question
  filter(UNDEMP == 1) %>% # only those wating to work additional hours
  filter(!is.na(ETH2)) %>%  # remove NAs from the ethnicity variable
  as_label(.)

```


```{r plot-longer-hours-eth, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.cap= "Ethnic-minority group are up to two times more likely to wish for more hours of work"}

p_undemp_eth  <- t_undemp_eth  %>% 
  ggplot(aes(x = sequence, y = percentage, colour = ETH2)) + 
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dark 3") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,0.50, 0.05), limits = c(0, 0.50), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "bottom",
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = "Note: Applies to all respondents who are not looking for a different or additional job"
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=0.50, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.35, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=0.50, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.35, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=0.50, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.35, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=.50, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.35, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06),
        legend.spacing = unit(0.2, 'cm')) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_undemp_eth

```


### Wanting more hours by qualification level

<p align="justify">Figure \@ref(fig:plot-longer-hours-quali) compares preferences for more hours in the current job by qualification. Overall, workers with lower level and no qualifications are substantially more likely to want to work more hours, certainly when compared with those with degrees who are again the group least likely to be underemployed.</p>

```{r longer-hours-quali-table, include=FALSE}

t_under_quali <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 2) %>% # Applies to all respondents who are not looking for a different or additional job
  group_by(year, sequence, quarter, quali, UNDEMP) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, quali) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(quali) & !quali %in% c("7")) %>% 
  filter(!is.na(UNDEMP)) %>% # remove NAs. Only responses to the question
  filter(UNDEMP == 1) %>% # only those wating to work additional hours
  as_label(.)

```

```{r plot-longer-hours-quali, fig.width = 7.5, fig.asp = 0.6, out.width = '100%', fig.cap= "The lower the qualification, the more likely to want more hours"}

p_under_quali <- t_under_quali %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(quali))) + 
  geom_line(size = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dark 3") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,0.3, 0.025), limits = c(0, 0.3), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), 
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title =" ",
       #subtitle  = "",
       caption = c("Note: Applies to all respondents who are not looking for a different or additional job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=0.3, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.25, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=0.3, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.25, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=0.30, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.25, label = "Covid-19", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=0.30, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.25, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
 guides(col = guide_legend(nrow = 2, byrow = TRUE, override.aes = list(size = 6)))

p_under_quali

```


### Wanting more hours by occupational group

<p align="justify"> That the risks of being underemployed are shaped by occupational group is also shown with this second indicator. Workers in routine and semi-routine occupations are more likely to report wanting to work more hours in their current jobs than are other workers (Figure \@ref(fig:plot-longer-hours-time-nsec)), peaking at almost one-fifth of the workers in 2013 and starting to rise again in 2022 after a post-pandemic dip.</p>

```{r longer-hours-table-NSEC, include=FALSE}

options(digits = 3)

t_under_nsec <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 2) %>% # Applies to all respondents who are not looking for a different or additional job
  group_by(year, sequence, quarter, NSEC2, UNDEMP) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, NSEC2) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(NSEC2) & !NSEC2 %in% c(3,6)) %>%  # omit no answers and no apply
  filter(!is.na(UNDEMP)) %>% # remove NAs. Only responses to the question
  filter(UNDEMP == 1) %>% # only those wanting to work additional hours
  as_label(.)

```


```{r plot-longer-hours-time-nsec, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Routine and semi-routine workers would like to work more hours in their current jobs"}

p_under_nsec <- t_under_nsec %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(NSEC2))) + 
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dynamic") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,0.5, 0.05), limits = c(0, 0.5), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), 
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = c("Note: Applies to all respondents who are not looking for a different or additional job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=0.5, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.40, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=0.5, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.40, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=0.5, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.40, label = "Covid-19", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=0.5, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.40, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_under_nsec


```


###  Wanting more hours by contract type

<p align="justify">When we examine who would like to work longer hours in their current jobs, we see again the higher proportions of those on zero hours contracts who are time-underemployed (Figure \@ref(fig:plot-longer-hours-fled)). This was especially true in the years following the great recession when around a third reported wanting to work more.</p>

```{r longer-hours-table-fled, include=FALSE}

options(digits = 3)

t_under_fled <- LFS_clean %>% # summary table
  filter(!sequence %in% c(1, 3, 5, 7, 9, 13, 17, 19, 21, 23, 
                          25, 27, 31, 33, 35, 37, 39, 41, 43,
                          45, 47, 49, 51, 53, 55)) %>% 
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 2) %>% # Applies to all respondents who are not looking for a different or additional job
  group_by(year, sequence, quarter, FLED, UNDEMP) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, FLED) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(FLED) & !FLED %in% c(5, 7, 8)) %>%  # omit compressed, none, and on call
  filter(!is.na(UNDEMP)) %>% # remove NAs. Only responses to the question
  filter(UNDEMP == 1) %>% # only those wanting to work additional hours
  as_label(.)
  
```

```{r plot-longer-hours-fled, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Workers in zero hours contracts more likely to want more hours of work"}

p_under_fled <- t_under_fled %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(FLED))) + 
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dynamic") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,1, 0.1), limits = c(0, 1), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), 
                     labels=c("","","2006","",
                              "","2007","","",
                              "","2008","","",
                              "","2009","","",
                              "","2010","","",
                              "","2011","","",
                              "","2012","","",
                              "","2013","","",
                              "","2014","","",
                              "","2015","","",
                              "","2016","","",
                              "","2017","","",
                              "","2018","","",
                              "","2019","","",
                              "","2020","","",
                              "","2021","","",
                              "","2022","",""),
                     expand = c(0,0)) +

   # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = c("Note: Applies to all respondents who are not looking for a different or additional job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=1, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.80, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=1, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.80, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.80, label = "Covid-19", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.80, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_under_fled

```


### Wanting more hours by part-time and full-time employment

<p align="justify">It is expected that those part-time workers who say they work part-time because they could not find full-time jobs would like to work more hours: Figure 1.15 shows that over half would like more hours. However, less is known about workers in full-time jobs and those workers who say that they are voluntarily working part-time. Are they fine with their current hours or would some also like more? Figure \@ref(fig:plot-longer-hours-fp-time) shows further that one in ten of those part-timers who work part-time because they do not want a full-time job would still like to work more hours. The ‘part-time’ hours category is a wide one, with shorter hours part-time jobs often associated with lower quality [@Warren2015]. It is notable that a smaller proportion of full-time workers express a desire for more hours compared to both groups of part-timers.</p> 


```{r extra-hours-ptimeftime-table, include=FALSE}

options(digits = 3)

t_undemp_ptime <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 2) %>% # Applies to all respondents who are not looking for a different or additional job
  group_by(year, sequence, quarter, FTPTWK, YPTJOB, UNDEMP) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, FTPTWK, YPTJOB) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(UNDEMP) & !is.na(FTPTWK)) %>% # remove NAs. Only responses to the question
  filter(UNDEMP == 1) %>%
  as_label(.)

t_undemp_ptime <- t_undemp_ptime %>% 
  mutate(YPTJOB = case_when(FTPTWK == "Full-time" ~ "Full-time", TRUE ~  YPTJOB)) %>% 
  filter(YPTJOB %in% c("Full-time", "Could not find full-time job", "Did not want full-time job"))

```

```{r plot-longer-hours-fp-time, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Some voluntary part-timers and full-time workers  would like to work more hours"}

p_undemp_rate_ptime <- t_undemp_ptime %>% 
  ggplot() + 
  geom_line(aes(x = sequence, y = percentage, colour = factor(YPTJOB)), size = 0.5, show.legend = FALSE) +  
  geom_point(aes(x = sequence, y = percentage, colour = factor(YPTJOB)), shape = 20, size = 1, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021
  #facet_wrap(. ~ SEX, ncol=1, strip.position = "top") +
  
  # scale and colour
  #scale_color_discrete_qualitative(palette = "Harmonic") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,1, 0.1), limits = c(0, 1), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "bottom",
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="Wants to work more hours in current job",
       #subtitle  = "Part-time vs full-time employed and self-employed workers",
       caption = c("Note: Applies to all respondents who are not looking for a different or additional job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=1, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.80, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=1, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.80, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.80, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.80, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +
  
  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_undemp_rate_ptime

```


<p align="justify">Knowing whether workers want to work more hours, or not, is a valuable indicator of time-related underemployment. It tells us little, however, about ‘how’ underemployed workers are, the intensity of their underemployment, and how far workers are away from achieving their hours preferences. Figure \@ref(fig:extra-hours-wanted-plot) shows that full and part-time workers who would like additional hours differ in the number of extra hours they would like to work. While full-time workers would like between 7 to 10 additional hours a week, those part-timers who could not find full-time would like up to 15 extra hours. Those who did not want full-time work would still like around 10 additional hours, with little change over time. Given that average hours worked by part-timers in the UK is 20 a week, this is evidence of persistent and wide actual/preferred hours gaps and deep time-related underemployment especially among part-timers.</p>

```{r table-extra-hours-pftime, include=FALSE}

t_extra_hours <- LFS_clean %>% 
  filter(AGE >= 18 & AGE <= 64) %>%
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 2) %>% # Applies to all respondents who are not looking for a different or additional job
  filter(UNDEMP == 1) %>%
  select(year, sequence, quarter, FTPTWK, YPTJOB, UNDHRS) %>%
  filter(!is.na(UNDHRS)) %>%
  as_label(FTPTWK, YPTJOB) %>%
  mutate(YPTJOB = case_when(FTPTWK == "Full-time" ~ "Full-time", TRUE ~  YPTJOB)) %>% 
  filter(YPTJOB %in% c("Full-time", "Could not find full-time job", "Did not want full-time job")) %>%
  group_by(year, sequence, quarter, YPTJOB) %>%
  summarise(median = median(UNDHRS)) 


```

```{r extra-hours-wanted-plot, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Extra hours wished to work by part-time and full-time workers"}

p_more_hours <- t_extra_hours %>% 
  ggplot(aes(x = sequence, y = median, colour = YPTJOB)) +
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(shape = 20, size = 1, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dark 3") + # Pastel 1, Dark 2, Dark 3, Set 2, Set 3, Warm, Cold, Harmonic, Dynamic
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,25, 5), limits = c(0, 25), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "bottom",
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Hours", 
       #title ="Median extra hours wished to work",
       #subtitle  = "Extra hours for those who could not find a full-time job",
       caption = c("Note: Applies to all respondents who are not looking for a different or additional job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=25, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 22, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=25, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 22, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=25, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 22, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=25, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 22, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +
  
  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_more_hours

```


### Wanting more hours by region

<p align="justify">All regions of the UK experienced similar levels of workers wanting to work more hours and all saw slight increases after economic recessions (Figure \@ref(fig:plot-more-hours-region)).</p>

```{r more-hours-region, include=FALSE}

LFS_clean <- LFS_clean %>% mutate(GOVTOF3 = case_match(GOVTOF2, 
                                            1 ~ "Rest of England",
                                            2 ~ "North West", 
                                            4 ~ "Rest of England",
                                            5 ~ "East Midlands", 
                                            6 ~ "Rest of England", 
                                            7 ~ "Rest of England",
                                            8 ~ "Rest of England", 
                                            9 ~ "Rest of England",
                                            10 ~ "South West",
                                            11 ~ "Wales",
                                            12 ~ "Scotland",
                                            13 ~ "Northern Ireland")) %>% 
                                                as_numeric(., keep.labels = TRUE)

LFS_clean$GOVTOF3 <- set_labels(LFS_clean$GOVTOF3, labels = c(
                                                      "East Midlands (incl. Nottingham)" = 1,
                                                      "North West (incl. Salford)" = 2, 
                                                      "Northern Ireland" = 3,
                                                      "Rest of England" = 4, 
                                                      "Scotland (incl. Glasgow)" = 5, 
                                                      "South West (incl. Bristol)" = 6,
                                                      "Wales" = 7))

options(digits = 2)

t_undemp_region <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 2) %>% # Applies to all respondents who are not looking for a different or additional job
  group_by(year, sequence, quarter, GOVTOF3, UNDEMP) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, GOVTOF3) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(GOVTOF3)) %>%  # omit no answers and no apply
  filter(!is.na(UNDEMP)) %>% # remove NAs. Only responses to the question
  filter(UNDEMP == 1) %>% # only those wating to work additional hours
  as_label(.)

```


```{r plot-more-hours-region, fig.width = 7, fig.asp = 0.6, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Slight increase in workers wanting extra hours after economic recessions and in all regions"}

p_undemp_region <- t_undemp_region %>% 
  filter(!GOVTOF3 %in% c("Northern Ireland","Wales")) %>% 
  ggplot(aes(x = sequence, y = percentage, col = factor(GOVTOF3))) + 
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dynamic") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,0.5, 0.05), limits = c(0, 0.5), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), 
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = c("Note: Applies to all respondents who are not looking for a different or additional job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=0.5, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.42, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=0.5, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.42, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=0.5, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.42, label = "Covid-19", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=0.5, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.42, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 2, byrow = TRUE, override.aes = list(size = 6)))

p_undemp_region

```


### Wanting more hours by industry

<p align="justify">Similar to the involuntary part-time indicator, Figure \@ref(fig:plot-more-hours-ind1) shows that the two sectors this project focuses on (Hospitality and Retail) have the highest proportion of workers wishing to work longer hours (followed by Education). The financial and information sectors again have low levels of underemployment, maintained over time, aside from a slight increase during the cost-of-living crisis.</p>

```{r more-hours-table-ind1, include=FALSE}

options(digits = 3)

t_undemp_ind1 <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 2) %>% # Applies to all respondents who are not looking for a different or additional job
  group_by(year, sequence, quarter, ind1, UNDEMP) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, ind1) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(ind1) & ind1 %in% c(7, 17, 11, 16, 10, 9)) %>%  # omit no answers and no apply
  filter(!is.na(UNDEMP)) %>% # remove NAs. Only responses to the question
  filter(UNDEMP == 1) %>% # only those wating to work additional hours
  as_label(.)

# Of interest:
  # "Wholesale and retail trade" = 7,
  # "Human health and social work" = 17

```


```{r plot-more-hours-ind1, fig.width = 7.5, fig.asp = 0.6, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Involuntary part-time is concentrated in the Retail and Hospitality sectors"}

p_undemp_ind1 <- t_undemp_ind1 %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(ind1))) + 
  geom_line(size = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dark 3") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,0.5, 0.05), limits = c(0, 0.5), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), 
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = c("Note: Applies to all respondents who are not looking for a different or additional job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=0.5, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.46, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=0.5, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.46, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=.5, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.46, label = "Covid-19", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=0.5, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.46, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 3, byrow = TRUE, override.aes = list(size = 6)))

p_undemp_ind1

```


## Time-underemployment indicator 3: Seeking a replacement job with more hours

<p align="justify">The third and final time-related underemployment indicator identifies those workers who are looking for a replacement job with more hours. For this, the Labour Force Survey asks participants whether they are looking for a different or additional paid job. If they answer yes, they are asked to clarify if it is an additional job or a replacement job. According to the definition of time-related underemployment, in this case we only consider those who are looking to replace their current job.</p> 

<p align="justify">Among those looking to change their job, between 15% to 20% would like to work longer hours in their new job. In addition, the percentage of employees in this situation increased after the Great Recession. However, since 2020 this trend has reverted, which could be linked to developments such as the UK pilot experience of reducing the work week to four days, as shown in Figure \@ref(fig:preferred-hours-replacement-job-sex).</p>


```{r preferred-job-hours-in-new-job-sex, include=FALSE}

t_seeking_hours <- LFS_clean %>% 
  filter(AGE >= 18 & AGE <= 64) %>%
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 1) %>%
  filter(ADDJOB == 1) %>%
  group_by(year, sequence, quarter, PREFHR) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(PREFHR) & PREFHR %in% c("1", "2", "3")) %>% 
  as_label(.)

```


```{r preferred-hours-replacement-job-sex, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Preferred working hours in the job being looked for"}

p_seeking_hours <- t_seeking_hours %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(PREFHR))) + 
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(shape = 20, size = 1, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021
  
  # scale and colour
  #scale_color_discrete_qualitative(palette = "Set 2") + # Pastel 1, Dark 2, Dark 3, Set 2, Set 3, Warm, Cold, Harmonic, Dynamic
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,1, 0.1), limits = c(0, 1), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "bottom",
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="Preferred working hours in new job being looked for",
       #subtitle  = " ",
       caption = c("Note: Applies to all workers seeking a new job to replace their present (main) job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=1, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.8, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=1, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.8, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.8, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = .8, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +
  
  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(2, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
 guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_seeking_hours

```


### Seeking a replacement job with more hours by sex

<p align="justify">Figure \@ref(fig:plot-seeking-job-sex) shows that between the period of 2006 to 2022 slightly more male than female workers sought a replacement job with more hours. Such differences reduced after the start of the cost of living crisis when female workers were more likely to look for a replacement job with more hours. This difference can be explained by the higher participation of men in the labour market compared to women, and the effects of the cost of living on female workers. </p>

```{r table-seeking-job-sex, include=FALSE}

#DIFJOB - Whether looking for different or additional paid job or business = (1) Yes
# ADDJOB - Whether seeking replacement or additional job = (1) A new job to replace your present (main) job

t_seeking <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 1) %>% # Whether looking for different or additional paid job or business = (1) Yes
  filter(ADDJOB == 1) %>% # replacement job 
  group_by(year, sequence, quarter, SEX, PREFHR) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, SEX) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(PREFHR == 1) %>% # more hours
  as_label(.)

```

```{r plot-seeking-job-sex, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.cap= "More male than female workers are seeking a replacement job"}

p_seeking_total <- t_seeking %>% 
  ggplot() + 
  geom_bar(aes(x = sequence, y = count, fill = SEX), stat="identity", position = "dodge", width=0.8) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_fill_discrete_qualitative(palette = "Dark 3") +
  scale_fill_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,500000, 50000), limits = c(0, 500000), labels = comma, expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "bottom", 
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "Year", 
       y = "Total", 
       #title ="Seeking replacement job",
       #subtitle  = "T ",
       caption = c("Note: Applies to all workers seeking a new job to replace their present (main) job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=500000, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 300000, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=500000, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 300000, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=500000, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 300000, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=500000, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 300000, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +
  
  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(2, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
 guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_seeking_total

```


### Seeking a replacement job with more hours by age group

<p align="justify">Considering the trends in the three indicators of time-related underemployment, younger workers could be considered more at risk of falling under this category than older workers. Specifically, Figure \@ref(fig:plot-seeking-by-age) shows that the majority of workers looking to replace their jobs are between 18 to 24 year old, followed by those between 25 to 34 years old.</p>

```{r seeking-by-age-table, include=FALSE}

options(digits = 3)

t_seeking_age <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  #creating age categories
  mutate(AGE_cat = case_when(AGE %in% 18:24 ~ "Aged 18 to 24",
                             AGE %in% 25:34 ~ "Aged 25 to 34",
                             AGE %in% 35:44 ~ "Aged 35 to 44",
                             AGE %in% 45:54 ~ "Aged 45 to 54",
                             AGE %in% 55:64 ~ "Aged 55 to 64",
                             is.na(AGE) ~ NA)) %>%
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 1) %>% # Whether looking for different or additional paid job or business = (1) Yes
  filter(ADDJOB == 1) %>% # replacement job 
  group_by(year, sequence, quarter, AGE_cat, PREFHR) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, AGE_cat) %>% # calculate % per group
  # percentage per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(PREFHR == 1) %>% # more hours
  as_label(.)

```


```{r plot-seeking-by-age, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, out.width = '100%', fig.cap= "Younger workers are more likely to look for a replacement job"}

p_seeking_age  <- t_seeking_age  %>% 
  ggplot(aes(x = sequence, y = count, fill = AGE_cat)) + 
  geom_bar(stat="identity", position = "stack", width=0.9) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_fill_discrete_qualitative(palette = "Dynamic") +
  scale_fill_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,500000, 50000), limits = c(0, 500000), labels = comma, expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "bottom", 
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Total", 
       #title ="Seeking replacement job",
       #subtitle  = " ",
       caption = c("Note: Applies to all workers seeking a new job to replace their present (main) job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=500000, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 450000, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=500000, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 450000, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=500000, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 450000, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=500000, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 450000, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +
  
  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(2, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
 guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_seeking_age
```


<p align="justify"> Furthermore, as shown in Figure \@ref(fig:preferred-hours-replacement-job-age) the group with the highest proportion of workers seeking to replace their jobs with one that has longer hours is concentrated among those between 18 and 24 years old.</p>

```{r preferred-job-hours-in-new-job-age, include=FALSE}

t_seeking_hours_age <- LFS_clean %>% 
  filter(AGE >= 18 & AGE <= 64) %>%
  #creating age categories
  mutate(AGE_cat = case_when(AGE %in% 18:24 ~ "Aged 18 to 24",
                             AGE %in% 25:34 ~ "Aged 25 to 34",
                             AGE %in% 35:44 ~ "Aged 35 to 44",
                             AGE %in% 45:54 ~ "Aged 45 to 54",
                             AGE %in% 55:64 ~ "Aged 55 to 64",
                             is.na(AGE) ~ NA)) %>%
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 1) %>%
  filter(ADDJOB == 1) %>%
  group_by(year, sequence, quarter, AGE_cat, PREFHR) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, AGE_cat) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(PREFHR) & PREFHR == 1) %>% 
  as_label(.)

```


```{r preferred-hours-replacement-job-age, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Workers between 18 to 24 years old looking for a replacement job with longer hours"}

p_seeking_hours_age <- t_seeking_hours_age %>% 
  ggplot(aes(x = sequence, y = percentage, colour = AGE_cat)) + 
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(shape = 20, size = 1, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021
  
  # scale and colour
  #scale_color_discrete_qualitative(palette = "Set 2") + # Pastel 1, Dark 2, Dark 3, Set 2, Set 3, Warm, Cold, Harmonic, Dynamic
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,1, 0.1), limits = c(0, 1), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "bottom",
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="Preferred working hours in new job being looked for",
       #subtitle  = " ",
       caption = c("Note: Applies to all workers seeking a new job to replace their present (main) job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=1, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.8, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=1, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.8, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.8, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = .8, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +
  
  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(2, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
 guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_seeking_hours_age

```


### Seeking a replacement job with more hours by ethnic group

<p align="justify"> As Figure \@ref(fig:plot-preferred-hours-replacement-job-eth) shows, the percentage of white workers looking to replace their jobs has consistently remained below 20% and relatively stable. In contrast, ethnic minority workers have consistently maintained percentages above  only dropping below 20% at the onset of the cost-of-living crisis.</p>

```{r table-preferred-job-hours-in-new-job-eth, include=FALSE}

LFS_clean <- LFS_clean %>% mutate(ETH2 = case_match(ETH, 
                                        1 ~ "White",
                                        2 ~ "Minority-ethnic", 
                                        3 ~ "Minority-ethnic", 
                                        4 ~ "Minority-ethnic", 
                                        5 ~ "Minority-ethnic",
                                        6 ~ "Minority-ethnic")) %>% 
                                    as_numeric(., keep.labels = TRUE)

t_seeking_hours_eth <- LFS_clean %>% 
  filter(AGE >= 18 & AGE <= 64) %>%
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 1) %>%
  filter(ADDJOB == 1) %>%
  group_by(year, sequence, quarter, ETH2, PREFHR) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, ETH2) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(PREFHR) & PREFHR == 1) %>% # only longer hours
  filter(!is.na(ETH2)) %>%  # remove NAs from the ethnicity variable
  as_label(.)

```


```{r plot-preferred-hours-replacement-job-eth, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Ethnic-minority workers more likely to look for replacement jobs with longer hours"}

p_seeking_hours_eth <- t_seeking_hours_eth %>% 
  ggplot(aes(x = sequence, y = percentage, colour = ETH2)) + 
  geom_line(linewidth = 0.7, show.legend = FALSE) +  
  geom_point(shape = 20, size = 1, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021
  
  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dark 3") + # Pastel 1, Dark 2, Dark 3, Set 2, Set 3, Warm, Cold, Harmonic, Dynamic
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,1, 0.1), limits = c(0, 1), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), position = "bottom",
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="Preferred working hours in new job being looked for",
       #subtitle  = " ",
       caption = c("Note: Applies to all workers seeking a new job to replace their present (main) job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=1, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.8, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=1, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.8, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.8, label = "Covid-19\nrecession", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = .8, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +
  
  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(2, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 4)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
 guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_seeking_hours_eth

```


### Seeking a replacement job with more hours by occupational group

<p align="justify">Lastly, workers in routine and semi-routine occupations are more interested in replacing their current job than are other workers. Figure \@ref(fig:plot-preferred-job-hours-in-new-job-nsec) shows furthermore that this group is considerably more likely to look for a replacement job with longer hours compared to management and professional occupations.</p>

```{r table-preferred-job-hours-in-new-job-NSEC, include=FALSE}

options(digits = 3)

t_seeking_hours_nsec <- LFS_clean %>% 
  filter(AGE >= 18 & AGE <= 64) %>%
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  filter(DIFJOB == 1) %>%
  filter(ADDJOB == 1) %>%
  group_by(year, sequence, quarter, NSEC2, PREFHR) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, NSEC2) %>% 
  # percentage part-time per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(!is.na(PREFHR) & PREFHR == 1) %>% # only longer hours
  filter(!is.na(NSEC2) & !NSEC2 %in% c("3","6")) %>%  # omit no answers and no apply
  as_label(.)

```

```{r plot-preferred-job-hours-in-new-job-nsec, fig.width = 7, fig.asp = 0.5, out.width = '100%', fig.fullwidth = TRUE, fig.cap= "Working class workers more likely to seek replacement jobs with longer hours"}

p_seeking_hours_nsec <- t_seeking_hours_nsec %>% 
  ggplot(aes(x = sequence, y = percentage, colour = factor(NSEC2))) + 
  geom_line(size = 0.7, show.legend = FALSE) +  
  geom_point(size = 1, shape = 20, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  #scale_color_discrete_qualitative(palette = "Dynamic") +
  scale_colour_okabeito(palette = "full") + 
  scale_y_continuous(breaks = seq(0,1, 0.1), limits = c(0, 1), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0)) +
  scale_x_continuous(breaks=c(1:68), 
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "", 
       y = "Percentage", 
       #title ="",
       #subtitle  = "",
       caption = c("Note: Applies to all workers seeking a new job to replace their present (main) job")
       ) +
  annotate('rect', xmin=10, xmax=14, ymin=0, ymax=1, alpha=.2, fill='brown1') + # Great Recession
  annotate(geom = "text", x = 10, y = 0.80, label = "Great Recession", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=30.5, xmax=31.5, ymin=0, ymax=1, alpha=.2, fill='darkgreen') + # back to pre-recession
  annotate(geom = "text", x = 31.2, y = 0.80, label = "UK economy gets back\nto pre-downturn size", hjust = 0, size = 5, colour = "darkgreen", lineheight = .7) +
  annotate('rect', xmin=56.5, xmax=58, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 56.5, y = 0.80, label = "Covid-19", hjust = 0, size = 5, colour = "darkred") +
  annotate('rect', xmin=63, xmax=68, ymin=0, ymax=1, alpha=.2, fill='brown1') +
  annotate(geom = "text", x = 63, y = 0.80, label = "Cost of\nliving crisis", hjust = 0, size = 5, colour = "darkred", lineheight = .7) +
  annotation_custom(copyright) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0.3), "cm"),
        plot.title = element_text(size = 15, margin = ggplot2::margin(0, 0, 0, 0)),
        plot.subtitle = element_text(size = 13, hjust = 1, margin = ggplot2::margin(0, 0, 4, 0), colour = "gray50"),
        plot.caption = element_text(size = 12, hjust = 1, margin = ggplot2::margin(0, 0, 2, 0), colour = "gray50"),
        axis.title.x = element_text(margin = ggplot2::margin(t = 10), size = 13),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 13),
        axis.text.x = element_text(size = 14, margin = margin(t = 2)),
        axis.text.y = element_text(size = 14),
        axis.text.y.left = element_text(margin = margin(r = 2)),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 15),
        panel.spacing = unit(.2, "cm"),
        panel.grid.major = element_line(color = "gray90", linewidth = .06)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 6)))

p_seeking_hours_nsec

```


## Summary: Time-related underemployment

<p align="justify">The time-related underemployed work fewer hours than they desire. In this report, we count as time-underemployed those part-timers who work part-time because they could not find a full-time job, workers (part- and full-timers) who would like to work longer hours in their current job, and workers seeking a replacement job with more hours. Approaching underemployment using three different measures gives us better insights into UK levels of this first form of underemployment and trends over time. We show which groups face higher levels of time-related underemployment and who is better protected from it. Results from the three indicators affirm that the groups facing higher levels of time-related underemployment include younger workers and workers from minority ethnic groups, as well as people employed in routine/semi-routine occupations and on precarious contracts. We move to skills-related underemployment next.</p>
