
```{r libraries-A1, include=FALSE, echo=FALSE}

#### packages needed
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(gridExtra)) install.packages("girdExtra", repos = "http://cran.us.r-project.org")
if(!require(Hmisc)) install.packages("Hmisc", repos = "http://cran.us.r-project.org")
if(!require(sjlabelled)) install.packages("sjlabelled", repos = "http://cran.us.r-project.org")
if(!require(ggthemes)) install.packages("ggthemes", repos = "http://cran.us.r-project.org")
if(!require(ggtext)) install.packages("ggtext", repos = "http://cran.us.r-project.org")
if(!require(ggrepel)) install.packages("ggrepel", repos = "http://cran.us.r-project.org")
if(!require(doBy)) install.packages("doBy", repos = "http://cran.us.r-project.org")
if(!require(sjmisc)) install.packages("sjmisc", repos = "http://cran.us.r-project.org")
if(!require(webshot)) install.packages("webshot", repos = "http://cran.us.r-project.org")
if(!require(scales)) install.packages("scales", repos = "http://cran.us.r-project.org")
if(!require(ggalt)) install.packages("ggalt", repos = "http://cran.us.r-project.org")
if(!require(janitor)) install.packages("janitor", repos = "http://cran.us.r-project.org")
if(!require(bookdown)) install.packages("bookdown", repos = "http://cran.us.r-project.org")
if(!require(htmlwidgets)) install.packages("htmlwidgets", repos = "http://cran.us.r-project.org")
if(!require(widgetframe)) install.packages("widgetframe", repos = "http://cran.us.r-project.org")
if(!require(officer)) install.packages("officer", repos = "http://cran.us.r-project.org")
if(!require(showtext)) install.packages("showtext", repos = "http://cran.us.r-project.org")
if(!require(colorspace)) install.packages("colorspace", repos = "http://cran.us.r-project.org")
if(!require(patchwork)) install.packages("patchwork", repos = "http://cran.us.r-project.org")
if(!require(ggpubr)) install.packages("ggpubr", repos = "http://cran.us.r-project.org")
if(!require(gt)) install.packages("gt", repos = "http://cran.us.r-project.org")
if(!require(htmltools)) install.packages("htmltools", repos = "http://cran.us.r-project.org")
if(!require(yaml)) install.packages("yaml", repos = "http://cran.us.r-project.org")

library(ggalt) # contingency table
library(tidyverse) # several tools for data manipulation
library(haven) # read SPSS, Stata, and SAS data
library(gridExtra) # arrange multiple grid-based plots on a page
library(Hmisc) # many functions useful for data analysis
library(sjmisc) # useful to set Na's
library(sjlabelled) # get_labels function
library(ggthemes) # themes for ggplot2
library(ggtext) # fancy text
library(ggrepel) # help with the position of the labels on ggplot2 graph
library(doBy) # 'do' something to data stratified 'by' some variables
library(scales) # add scale symbols and colour to axis
library(janitor) # add totals to tables
library(bookdown) # add totals to tables
library(htmlwidgets)
library(widgetframe)
library(colorspace) # colour blind friendly pallettes 
library(officer) # to add additional formatting to tables 
library(patchwork) # to put graphs side by side
library(ggpubr) # easy-to-use functions for creating and customizing ‘ggplot2’- based publication ready plots.
library(gt) #  creation of interactive data tables
library(htmltools) # tools for creating, manipulating, and writing HTML 
library(webshot)
library(yaml)
library(showtext)
  font_add_google("Lato")
  showtext_auto()

```

# Technical Appendix 1: The Data {-} 

<p align="justify">We use the UK’s largest study on employment circumstances, the Labour Force Survey (LFS), drawing on analysis from the 2006 to 2022 releases (calendar quarters). We take a pseudo-longitudinal approach, where we analyse repeated cross-sectional data (Table 1). It is important to note the following:</p>
-	<p align="justify">We calculate the underemployment rate based on the employed population rather than the total active population. Between wave 2 (April-June) of 2020 and wave 1 (January-March) of 2022, sample sizes were significantly affected by the Covid-19 pandemic, potentially affecting results from this period.</p>
-	<p align="justify">From wave 2 (April-June) in 2011, the ethnicity variable was modified in accordance to Census data. In this report we have recorded the variable to ensure consistency across all quarters in the dataset.</p>
-	<p align="justify">The job arrangements have also been recorded to ensure consistency across all quarters in the dataset, as this variable was modified three times during the periods included in this analysis.</p>

```{r sample, include=FALSE, echo=FALSE}

load("../../data/LFS_clean.rda")

N_total <- LFS_clean %>% 
  group_by(as_factor(year), 
           as_label(quarter)) %>%
  summarise(Active = sum(na.omit(PWT[ILODEFR <=2 & AGE >= 18 & AGE <= 64])), # in employment and unemployed between 18 to 64
            Workers = sum(na.omit(PWT[ILODEFR == 1 & INECAC05 <= 2 & AGE >= 18 & AGE <= 64])),
            Women_workers = sum(na.omit(PWT[ILODEFR == 1 & INECAC05 <= 2 & SEX == 2 & AGE >= 18 & AGE <= 64])),
            Men_workers = sum(na.omit(PWT[ILODEFR == 1 & INECAC05 <= 2 & SEX == 1 & AGE >= 18 & AGE <= 64]))) %>% # employee and self-employees
  mutate(across(c(Active, Workers, Women_workers, Men_workers), comma)) %>%
  rename(Year = `as_factor(year)`,
         Quarter = `as_label(quarter)`,
         "Total active" = Active,
         "Total in employment" = Workers,
         "Female workers" = Women_workers,
         "Male workers" = Men_workers) %>% # Total population estimates
  group_by(Year)

```

```{r table-html, echo=FALSE}

N_total %>% gt(groupname_col = NULL) %>% 
   tab_header(
    title = md("**Table 1: Calendar quarters from 2006 to 2022**"),
    subtitle = md("Population weights are used.")) %>% 
  tab_footnote(footnote = "Active population includes employed and unemployed persons",
               locations = cells_column_labels(columns = "Total active")) %>%
  tab_source_note(source_note = 'Source: UK Labour Force Survey.') %>%
  tab_options(ihtml.active = TRUE,
              ihtml.use_pagination = TRUE,
              ihtml.page_size_default = 4,
              heading.title.font.size = 15,
              footnotes.multiline = TRUE,
              source_notes.multiline = TRUE)
    
```

