---
title: "Measuring Underemployment - Wages"
subtitle: 'Trends from 2006 to 2022'
author: "The Underemployment Project"
date: '`r Sys.Date()`'

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
	warning = FALSE,
  embed_fonts = TRUE,
  fig.align = "center",
  fig_caption= TRUE,
  use_bookdown= TRUE,
  toc= FALSE,
  toc_depth= 3
)
```

```{r libraries, include=FALSE}

#### packages needed
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(gridExtra)) install.packages("girdExtra", repos = "http://cran.us.r-project.org")
if(!require(Hmisc)) install.packages("Hmisc", repos = "http://cran.us.r-project.org")
if(!require(sjlabelled)) install.packages("sjlabelled", repos = "http://cran.us.r-project.org")
if(!require(ggthemes)) install.packages("ggthemes", repos = "http://cran.us.r-project.org")
if(!require(ggtext)) install.packages("ggtext", repos = "http://cran.us.r-project.org")
if(!require(ggrepel)) install.packages("ggrepel", repos = "http://cran.us.r-project.org")
if(!require(doBy)) install.packages("doBy", repos = "http://cran.us.r-project.org")
if(!require(sjmisc)) install.packages("sjmisc", repos = "http://cran.us.r-project.org")
if(!require(flextable)) install.packages("flextable", repos = "http://cran.us.r-project.org")
if(!require(scales)) install.packages("scales", repos = "http://cran.us.r-project.org")
if(!require(ggalt)) install.packages("ggalt", repos = "http://cran.us.r-project.org")
if(!require(janitor)) install.packages("janitor", repos = "http://cran.us.r-project.org")
if(!require(bookdown)) install.packages("bookdown", repos = "http://cran.us.r-project.org")
if(!require(rmdformats)) install.packages("rmdformats", repos = "http://cran.us.r-project.org")
if(!require(officer)) install.packages("officer", repos = "http://cran.us.r-project.org")
if(!require(showtext)) install.packages("showtext", repos = "http://cran.us.r-project.org")
if(!require(colorspace)) install.packages("colorspace", repos = "http://cran.us.r-project.org")
if(!require(patchwork)) install.packages("patchwork", repos = "http://cran.us.r-project.org")
if(!require(sjstats)) install.packages("sjstats", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")

library(ggalt) # contingency table
library(tidyverse) # several tools for data manipulation
library(haven) # read SPSS, Stata, and SAS data
library(gridExtra) # arrange multiple grid-based plots on a page
library(Hmisc) # many functions useful for data analysis
library(sjmisc) # useful to set Na's
library(sjlabelled) # get_labels function
library(ggthemes) # themes for ggplot2
library(ggtext) # fancy text
library(ggrepel) # help with the position of the labels on ggplot2 graph
library(doBy) # 'do' something to data stratified 'by' some variables
library(flextable) # formatting tables 
library(scales) # add scale symbols and colour to axis
library(janitor) # add totals to tables
library(rmdformats)
library(sjstats) # add weights 
library(colorspace) # colour blind friendly pallettes 
library(officer) # to add additional formatting to tables 
library(patchwork) # to put graphs side by side
library(caret) # machine learning
library(showtext)
  font_add_google("Lato")
  showtext_auto()


```

```{r load data, include=FALSE}

load("../data/LFS_clean.rda")

```



## Background

This report aims at mapping underemployment trends for employees and self-employed workers in the UK. We use the UK’s largest study on employment circumstances, the Labour Force Survey (LFS), drawing on analysis from the 2006 to 2022 releases (calendar quarters). Our analyses consider only those in employment  between 18 and 64 years old.

```{r sample, echo=FALSE}

N_total <- LFS_clean %>% 
  group_by(as_factor(year), 
           as_label(quarter)) %>%
  summarise(Population = sum(na.omit(PWT[AGE >= 18 & AGE <= 64])), # everyone between 18 to 64
            Active = sum(PWT[ILODEFR <=2 & AGE >= 18 & AGE <= 64]), # in employment and unemployed between 18 to 64
            Workers = sum(PWT[ILODEFR == 1 & INECAC05 <= 2 & AGE >= 18 & AGE <= 64])) %>% # employee and self-employees between 18 to 64
  rename(Year = `as_factor(year)`,
         Quarter = `as_label(quarter)`) # Total population estimates
  
```


In this bulletin, low and high pay are calculated using the Organisation for Economic Co-operation and Development (OECD) definitions . Low pay is defined as the value that is two-thirds of median hourly earnings and high pay is defined as the value that is 1.5 times median hourly earnings.


```{r low earners table}

options(digits = 3)

t_low_earners <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c("1")) %>% #only employees
  filter(HOURPAY > 0 & SUMHRS > 0) %>%
  group_by(year, sequence, quarter) %>%
  summarise(median_pay = weighted_median(HOURPAY, weights = PIWT),
            low_pay = 2*(weighted_median(HOURPAY, weights = PIWT))/3,
            low_earner_men_median = sum(PIWT[SEX == 1 & HOURPAY <= low_pay]),
            low_earner_women_median = sum(PIWT[SEX == 2 & HOURPAY <= low_pay])) %>%
  as_label(.) %>%
  rename(Men = low_earner_men_median,
         Women = low_earner_women_median) %>%
  pivot_longer(c(Men, Women), names_to = "sex", values_to = "total")
  

```

```{r low earners plot, fig.width= 9, fig.height=3.5, fig.fullwidth=TRUE, fig.cap= ""}

p_low_earners  <- t_low_earners %>% 
  ggplot() + 
  geom_bar(aes(x = sequence, y = total, fill = sex), stat="identity", position = "dodge", width=0.8) +  
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  scale_fill_discrete_qualitative(palette = "Dark 3") +
  scale_y_continuous(breaks = seq(0,4000000, 400000), limits = c(0, 4000000), labels = comma, expand = c(0,0), sec.axis = dup_axis()) +
  scale_x_continuous(breaks=c(1:68),
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "Year", 
       y = "Total", 
       title ="Low pay employees",
       subtitle  = "Employees paid less than two-thirds of median hourly pay",
       #caption = c("Note: Only workers who are not looking for a different or additional job")
       ) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0), "lines"),
        plot.title = element_text(size = 14, margin = ggplot2::margin(5, 0, 0, 0)),
        plot.subtitle = element_text(size = 12, hjust = 0, margin = ggplot2::margin(5, 0, 2, 0)),
        plot.caption = element_text(size = 10, hjust = 1, margin = ggplot2::margin(2, 0, 2, 0)),
        axis.title.x = element_text(margin = ggplot2::margin(t = 8), size = 12),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 12),
        axis.text.x = element_text(size = 12, margin = margin(t = 2)),
        axis.text.y = element_text(size = 12),
        axis.text.y.right =  element_text(margin = margin(l = 4)),
        axis.title.y.right = element_blank(),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.grid.major = element_line(color = "gray90", linewidth = .5)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 3)))

p_low_earners

```



```{r}

set.seed(300, sample.kind="Rounding")

data_ptime <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1)) %>% #only employees
  filter(FTPTWK == 2) %>% # only part-timers
  filter(YPTJOB %in% c(3, 4)) %>%
  filter(!is.na(YPTJOB)) %>%
  filter(HOURPAY > 0 & SUMHRS > 0) %>%
  filter(!is.na(HOURPAY))


fit_rtree <- train(factor(YPTJOB) ~ HOURPAY, data = data_ptime,
                   method = "rpart",
                   tuneGrid = data.frame(cp = seq(0, 0.05, len = 25)))

ggplot(fit_rtree, highlight = TRUE)


```






