---
title: "Measuring Underemployment"
subtitle: 'Trends from 2006 to 2022'
author: "The Underemployment Project"
date: '`r Sys.Date()`'

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
	warning = FALSE,
  embed_fonts = TRUE,
  fig.align = "center",
  fig_caption= TRUE,
  use_bookdown= TRUE,
  toc= FALSE,
  toc_depth= 3
)
```

```{r libraries, include=FALSE}

#### packages needed
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(gridExtra)) install.packages("girdExtra", repos = "http://cran.us.r-project.org")
if(!require(Hmisc)) install.packages("Hmisc", repos = "http://cran.us.r-project.org")
if(!require(sjlabelled)) install.packages("sjlabelled", repos = "http://cran.us.r-project.org")
if(!require(ggthemes)) install.packages("ggthemes", repos = "http://cran.us.r-project.org")
if(!require(ggtext)) install.packages("ggtext", repos = "http://cran.us.r-project.org")
if(!require(ggrepel)) install.packages("ggrepel", repos = "http://cran.us.r-project.org")
if(!require(doBy)) install.packages("doBy", repos = "http://cran.us.r-project.org")
if(!require(sjmisc)) install.packages("sjmisc", repos = "http://cran.us.r-project.org")
if(!require(plotly)) install.packages("plotly", repos = "http://cran.us.r-project.org")
if(!require(flextable)) install.packages("flextable", repos = "http://cran.us.r-project.org")
if(!require(scales)) install.packages("scales", repos = "http://cran.us.r-project.org")
if(!require(ggalt)) install.packages("ggalt", repos = "http://cran.us.r-project.org")
if(!require(janitor)) install.packages("janitor", repos = "http://cran.us.r-project.org")
if(!require(bookdown)) install.packages("bookdown", repos = "http://cran.us.r-project.org")
if(!require(rmdformats)) install.packages("rmdformats", repos = "http://cran.us.r-project.org")
if(!require(htmlwidgets)) install.packages("htmlwidgets", repos = "http://cran.us.r-project.org")
if(!require(widgetframe)) install.packages("widgetframe", repos = "http://cran.us.r-project.org")
if(!require(officer)) install.packages("officer", repos = "http://cran.us.r-project.org")
if(!require(showtext)) install.packages("showtext", repos = "http://cran.us.r-project.org")
if(!require(colorspace)) install.packages("colorspace", repos = "http://cran.us.r-project.org")
if(!require(patchwork)) install.packages("patchwork", repos = "http://cran.us.r-project.org")
if(!require(ggpubr)) install.packages("ggpubr", repos = "http://cran.us.r-project.org")

library(ggalt) # contingency table
library(tidyverse) # several tools for data manipulation
library(haven) # read SPSS, Stata, and SAS data
library(gridExtra) # arrange multiple grid-based plots on a page
library(Hmisc) # many functions useful for data analysis
library(sjmisc) # useful to set Na's
library(sjlabelled) # get_labels function
library(ggthemes) # themes for ggplot2
library(ggtext) # fancy text
library(ggrepel) # help with the position of the labels on ggplot2 graph
library(doBy) # 'do' something to data stratified 'by' some variables
library(plotly) # interactive plots
library(flextable) # formatting tables 
library(scales) # add scale symbols and colour to axis
library(janitor) # add totals to tables
library(rmdformats)
library(htmlwidgets)
library(widgetframe)
library(colorspace) # colour blind friendly pallettes 
library(officer) # to add additional formatting to tables 
library(patchwork) # to put graphs side by side
library(ggpubr)
library(showtext)
  font_add_google("Lato")
  showtext_auto()


```

```{r load data, include=FALSE}

load("../data/LFS_clean.rda")

```



## Background

<p align="justify">This report aims at mapping underemployment trends for employees and self-employed workers in the UK. We use the UK’s largest study on employment circumstances, the Labour Force Survey (LFS), drawing on analysis from the 2006 to 2022 releases (calendar quarters). Our analyses consider only those in employment  between 18 and 64 years old.</p>


```{r sample, echo=FALSE}

N_total <- LFS_clean %>% 
  group_by(as_factor(year), 
           as_label(quarter)) %>%
  summarise(Population = sum(na.omit(PWT[AGE >= 18 & AGE <= 64])), # everyone between 18 to 64
            Active = sum(PWT[ILODEFR <=2 & AGE >= 18 & AGE <= 64]), # in employment and unemployed between 18 to 64
            Workers = sum(PWT[ILODEFR == 1 & INECAC05 <= 2 & AGE >= 18 & AGE <= 64])) %>% # employee and self-employees between 18 to 64
  rename(Year = `as_factor(year)`,
         Quarter = `as_label(quarter)`) # Total population estimates
  

```

## Skills mismatch

<p align="justify"> We follow definitions the normative and statistical approaches of skills mismatch established by ILO and ONS to calculate:
■ people who are overeducated for their occupation. This, based on the normative approach of matching an individual's higher qualification (ISCDE) and occupation (ISCO) first-digit indicator.
■ people who are overeducated for their occupation. This, based on the statistical approach of matching an individual's higher qualification dispersion (ISCDE) and occupation (ISCO) three-digit indicator.</p>


### normative approach to skills mismatch

```{r normative approach to skills mismatch table, echo=FALSE}


options(digits = 3)

v_skills <- LFS_clean %>% # summary table
  filter(AGE >= 18 & AGE <= 64) %>% 
  filter(ILODEFR == 1) %>% # only people in employment
  filter(INECAC05 %in% c(1, 2)) %>% #only employees and self-employed
  group_by(year, sequence, quarter, SEX, sk_mm) %>% 
  summarise(count = sum(PWT)) %>% 
  group_by(year, sequence, quarter, SEX) %>% 
  # percentage per quarter
  mutate(percentage = count/sum(count)) %>% 
  filter(sk_mm == 1) %>%  #filtering by overeducated
  as_label(.)

```


```{r plot skills overeducated employees, fig.width=9, fig.height=6, fig.fullwidth=TRUE, fig.cap= "workers who have a qualification above the required for their occupation"}

p_nskills <- n_skills %>% 
  ggplot() + 
  geom_line(aes(x = sequence, y = percentage, colour = factor(SEX)), size = 0.5, show.legend = FALSE) +  
  geom_point(aes(x = sequence, y = percentage, colour = factor(SEX)), shape = 15, show.legend = TRUE) +
  geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021

  # scale and colour
  scale_color_discrete_qualitative(palette = "Dark 3") +
  scale_y_continuous(breaks = seq(0,1, 0.1), limits = c(0, 1), labels =
                       scales::percent_format(accuracy = 1L), expand = c(0,0), sec.axis = dup_axis()) +
  scale_x_continuous(breaks=c(1:68), position = "top",
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  # annotation
  labs(x = "Year", 
       y = "Percentage", 
       title ="Figure 1: Overeducated workers",
       subtitle  = "",
       caption = ""
       ) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0), "lines"),
        plot.title = element_text(size = 14, margin = ggplot2::margin(5, 0, 0, 0)),
        plot.subtitle = element_text(size = 12, hjust = 0, margin = ggplot2::margin(5, 0, 15, 0)),
        plot.caption = element_text(size = 10, hjust = 1, margin = ggplot2::margin(2, 0, 2, 0)),
        axis.title.x = element_text(margin = ggplot2::margin(t = 8), size = 12),
        axis.title.y = element_text(margin = ggplot2::margin(r = 5), size = 12),
        axis.text.x = element_text(size = 12, margin = margin(t = 2)),
        axis.text.y = element_text(size = 12),
        axis.text.y.right =  element_text(margin = margin(l = 4)),
        axis.title.y.right = element_blank(),
        axis.ticks.length.x = unit(0.1, "cm"), 
        legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.grid.major = element_line(color = "gray90", linewidth = .5)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 5)))

nskills_total <- v_skills %>% 
  ggplot() + 
  geom_bar(aes(x = sequence, y = count, fill = factor(SEX)), stat = "identity", position = "dodge", width=0.9) +   geom_vline(xintercept = 4.5, color = "#c9c9c9") + # gridline 2006
  geom_vline(xintercept = 8.5, color = "#c9c9c9") + # gridline 2007
  geom_vline(xintercept = 12.5, color = "#c9c9c9") + # gridline 2008
  geom_vline(xintercept = 16.5, color = "#c9c9c9") + # gridline 2009
  geom_vline(xintercept = 20.5, color = "#c9c9c9") + # gridline 2010
  geom_vline(xintercept = 24.5, color = "#c9c9c9") + # gridline 2011
  geom_vline(xintercept = 28.5, color = "#c9c9c9") + # gridline 2012
  geom_vline(xintercept = 32.5, color = "#c9c9c9") + # gridline 2013
  geom_vline(xintercept = 36.5, color = "#c9c9c9") + # gridline 2014
  geom_vline(xintercept = 40.5, color = "#c9c9c9") + # gridline 2015
  geom_vline(xintercept = 44.5, color = "#c9c9c9") + # gridline 2016
  geom_vline(xintercept = 48.5, color = "#c9c9c9") + # gridline 2017
  geom_vline(xintercept = 52.5, color = "#c9c9c9") + # gridline 2018
  geom_vline(xintercept = 56.5, color = "#c9c9c9") + # gridline 2019
  geom_vline(xintercept = 60.5, color = "#c9c9c9") + # gridline 2020
  geom_vline(xintercept = 64.5, color = "#c9c9c9") + # gridline 2021


  # scale and colour
  scale_fill_discrete_qualitative(palette = "Dark 3") +
  scale_x_continuous(breaks=c(1:68),
                     labels=c("","2006","", "", "", "2007", "", "", "", "2008", "", "", "", "2009", "", 
                            "", "", "2010", "", "", "", "2011", "", "", "", "2012", "", "", 
                            "", "2013", "", "", "", "2014", "", "", "", "2015", "", "", "", 
                            "2016", "", "", "", "2017", "", "", "", "2018", "", "", "", "2019", "",
                            "", "", "2020", "", "", "", "2021", "", "", "", "2022", "", ""), expand = c(0,0)) +
  scale_y_continuous(breaks = seq(0,10000000,1000000), limits = c(0,10000000), labels = comma, expand = c(0,0), sec.axis = dup_axis()) +
  
  # annotation
  labs(x = "Year", 
       y = "Total", 
       title = "Figure 2: Overeducated workers (total)",
       #subtitle  = "Totals based on people in employment between 2006 and 2022",
       #caption = c("Note: People in employment (employees and self-employed) between 2006 and 2022")
       ) +

  # theme
  theme_economist_white(gray_bg = FALSE, base_family = "Lato") +
  theme(plot.margin = unit(c(0, 0, 0, 0), "lines"),
        plot.title = element_text(size = 14, margin = ggplot2::margin(5, 0, 0, 0)),
        plot.subtitle = element_text(size = 12, hjust = 0, margin = ggplot2::margin(5, 0, 2, 0)),
        plot.caption = element_text(size = 10, hjust = 1, margin = ggplot2::margin(2, 0, 2, 0)),
        axis.title.x = element_text(margin = ggplot2::margin(t = 8), size = 12),
        axis.title.y = element_text(margin = ggplot2::margin(r = 2), size = 12),
        axis.text.x = element_text(size = 12, margin = margin(t = 2)),
        axis.text.y = element_text(size = 12),
        axis.ticks.length.x = unit(0.1, "cm"), 
        axis.text.y.right =  element_text(margin = margin(l = 4)),
        axis.title.y.right = element_blank(),
        legend.position = "none", 
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.grid.major = element_line(color = "gray90", linewidth = .5)) +
  guides(col = guide_legend(nrow = 1, byrow = TRUE, override.aes = list(size = 5)))


p_vskills / p_vskills_total



```


## Feedback PG

### disability 
### elements around poverty 
### caring responsibilities
### Green jobs 


## Feedback AG

### Wages and employment conditions could be predictors of underemployment (time and skills)
### Overwork 
### Check working conditions 
### Geography is important to consider (e.g., rural vs urban)
### People may want to work more hours because they are poorly paid
### Living wages 
### CIPD survey is useful for a subjective underemployment... we could add some other questions there.
### Check for groups (like students)
### Language suggestions: workforce planning, simplidy the concepts (low pay, use of skills, extra hours)

