---
title: "Measuring Underemployment"
subtitle: 'Trends from 2006 to 2022'
author: ""
date: '`r Sys.Date()`'
output: 
  word_document:
      reference_docx: "Report_template.docx"
      toc: FALSE
      toc_depth: 3

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
	warning = FALSE,
  collapse = TRUE)
```

```{r libraries, include=FALSE}

#### packages needed
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(gridExtra)) install.packages("girdExtra", repos = "http://cran.us.r-project.org")
if(!require(Hmisc)) install.packages("Hmisc", repos = "http://cran.us.r-project.org")
if(!require(sjlabelled)) install.packages("sjlabelled", repos = "http://cran.us.r-project.org")
if(!require(ggthemes)) install.packages("ggthemes", repos = "http://cran.us.r-project.org")
if(!require(ggtext)) install.packages("ggtext", repos = "http://cran.us.r-project.org")
if(!require(ggrepel)) install.packages("ggrepel", repos = "http://cran.us.r-project.org")
if(!require(doBy)) install.packages("doBy", repos = "http://cran.us.r-project.org")
if(!require(sjmisc)) install.packages("sjmisc", repos = "http://cran.us.r-project.org")
if(!require(plotly)) install.packages("plotly", repos = "http://cran.us.r-project.org")
if(!require(flextable)) install.packages("flextable", repos = "http://cran.us.r-project.org")
if(!require(scales)) install.packages("scales", repos = "http://cran.us.r-project.org")
if(!require(ggalt)) install.packages("ggalt", repos = "http://cran.us.r-project.org")
if(!require(janitor)) install.packages("janitor", repos = "http://cran.us.r-project.org")
if(!require(bookdown)) install.packages("bookdown", repos = "http://cran.us.r-project.org")
if(!require(rmdformats)) install.packages("rmdformats", repos = "http://cran.us.r-project.org")
if(!require(htmlwidgets)) install.packages("htmlwidgets", repos = "http://cran.us.r-project.org")
if(!require(widgetframe)) install.packages("widgetframe", repos = "http://cran.us.r-project.org")
if(!require(officer)) install.packages("officer", repos = "http://cran.us.r-project.org")
if(!require(showtext)) install.packages("showtext", repos = "http://cran.us.r-project.org")
if(!require(colorspace)) install.packages("colorspace", repos = "http://cran.us.r-project.org")

library(ggalt) # contingency table
library(tidyverse) # several tools for data manipulation
library(haven) # read SPSS, Stata, and SAS data
library(gridExtra) # arrange multiple grid-based plots on a page
library(Hmisc) # many functions useful for data analysis
library(sjmisc) # useful to set Na's
library(sjlabelled) # get_labels function
library(ggthemes) # themes for ggplot2
library(ggtext) # fancy text
library(ggrepel) # help with the position of the labels on ggplot2 graph
library(doBy) # 'do' something to data stratified 'by' some variables
library(plotly) # interactive plots
library(flextable) # formatting tables 
library(scales) # add scale symbols and colour to axis
library(janitor) # add totals to tables
library(rmdformats)
library(htmlwidgets)
library(widgetframe)
library(colorspace) # colour blind friendly pallettes 
library(officer) # to add additional formatting to tables 
library(showtext)
  font_add_google("Lato")
  showtext_auto()


```

```{r load data, include=FALSE}

load("../data/LFS_clean.rda")

```



## Background



### By hours

## People in part-time work who wants to work more hours

# Index 1
### FTPTWK 2 = part time    
### YPTJOB 3 = could not find
### DIFJOB, # Whether looking for different or additional paid job or business
### UNDEMP", # Whether would like to work longer hours, at current basic rate of pay (filter by DIFJOB = 2)



```{r index 1, echo=FALSE}





```



```{r index 1 plot, fig.width= 8, fig.height=5}



```

# Index 2
### FTPTWK 2 = part time    
### YPTJOB 3 = could not find
### DIFJOB, # Whether looking for different or additional paid job or business
### UNDHRS, # Number of extra hours would like to work (filter by UNDEMP = 1)


```{r index 2, echo=FALSE}



```



```{r index 2 plot, fig.width= 8, fig.height=5}



```




# Income

### 1.25 times the poverty threshold (international and UK definition)

```{r}

```



```{r}

```




